---
title: "NBA_Player_sel_analysis"
output: html_document
---
### Introduction

Normalisescoring rates by standardising to score per 40 min (Kubatko, 2007)
several models have been attempted (Spector 2020). difficult to predict how 5 opponents will perform WITH each otehr and also against an opposition
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
```

## Reading and Cleaning the Raw Data

```{r read in data, echo=FALSE, include=FALSE}
df_sal <- read_csv("data_raw/2018-19_nba_player-salaries_raw.csv")
df_players <- read_csv("data_raw/2018-19_nba_player-statistics_raw.csv")
df_team_payroll <- read_csv("data_raw/2019-20_nba_team-payroll_raw.csv")
df_team_ptsMP <- read_csv("data_raw/2018-19_nba_team-statistics_2_raw.csv")
df_team_WL <- read_csv("data_raw/2018-19_nba_team-statistics_1_raw.csv")
```

```{r convert payroll(cha) to num and scale, include=FALSE}
df_team_payroll$COL1 <- as.numeric(gsub('[$,]', '', df_team_payroll$salary))
df_team_payroll <- mutate(df_team_payroll, COL2 = COL1/1000000)
```

Add standardised team scoring rate per 40min period (Kubatko, 2007), and win%
```{r add column, include=FALSE}
df_team_pts40 <- mutate(df_team_ptsMP, team_pts40 = (PTS / MP)* 40 )
```


[This is a link to the data fields](https://uclearn.canberra.edu.au/courses/9531/pages/data-description-reproducible-data-analysis-project)

**Chicago Bulls payroll is 24th or the 30 NBA teams**
```{r}
# Barplot team payrolls
team_sal <- ggplot(data = df_team_payroll, aes(x = reorder(team, COL2), y = COL2)) +
  geom_bar(stat="identity")
# Horizontal bar plot
new <- team_sal + coord_flip(ylim = c(50, 160))
#label axes
new + labs(title = "Team payroll (2018-19)", face = "italics", x = "Team", y = "Millions") 
```
**Win % versus team salary**
```{r}
df_team_WL <- mutate(df_team_WL, winP = (W / (W + L)* 100))
#add win% column
df_team_pts40 <- bind_cols(df_team_pts40, df_team_WL[-c(1:25)])
# team average pts per minute
team_pts40 <- ggplot(data = df_team_pts40, aes(x = reorder(Team, winP), y = winP)) +
  geom_bar(stat="identity")
# Horizontal bar plot
new1 <- team_pts40 + coord_flip(ylim = c(20, 75)) +
  labs(title = "Team game win percentages (2018-19)", x = "Team", y = "Win %")
new1
```



**Chicago Bulls are ranked 29th of 30 in season average points scored per minute played**
```{r barplot}
# team average pts per minute
ggteam_pts40 <- ggplot(data = df_team_pts40, aes(x = reorder(Team, team_pts40), y = team_pts40)) +
  geom_bar(stat="identity")
# Horizontal bar plot
new2 <- ggteam_pts40 + coord_flip(ylim = c(17, 20)) +
  labs(title = "Team average points scored per 40 minutes played (2018-19)", x = "Team", y = "Points / 40 Minutes")
new2
```
*standardise variables to events per 40min*
```{r establish new variables}
df_team_pts40 <- df_team_pts40 %>%
  mutate(AST40 = (AST / MP)* 40,
         TOV40 = (TOV / MP)* 40,
         STL40 = (STL / MP)* 40,
         BLK40 = (BLK / MP)* 40,
         PF40 = (PF / MP)* 40,
         TRB40 = (TRB / MP)* 40,
         ORB40 = (ORB / MP)* 40,
         DRB40 = (DRB / MP)* 40,
         FG40 = (FG / MP)* 40,
         FGA40 = (FGA / MP)* 40)
```


**correlation wins and points scored**
```{r}

#points/40 vs win%
df_team_pts40 %>%
  ggplot(aes(x = team_pts40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$team_pts40, y = df_team_pts40$winP, method = "pearson")

#assists vs win%
df_team_pts40 %>%
  ggplot(aes(x = AST40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$AST40, y = df_team_pts40$winP, method = "pearson")

#turnovers vs win%
df_team_pts40 %>%
  ggplot(aes(x = TOV40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$TOV40, y = df_team_pts40$winP, method = "pearson")

#steals vs win%
df_team_pts40 %>%
  ggplot(aes(x = STL40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$STL40, y = df_team_pts40$winP, method = "pearson")

#blocks vs win%
df_team_pts40 %>%
  ggplot(aes(x = BLK40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$BLK40, y = df_team_pts40$winP, method = "pearson")

#fouls vs win%
df_team_pts40 %>%
  ggplot(aes(x = PF40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$PF40, y = df_team_pts40$winP, method = "pearson")

#TRB vs win%
df_team_pts40 %>%
  ggplot(aes(x = TRB40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$TRB40, y = df_team_pts40$winP, method = "pearson")

#ORB vs win%
df_team_pts40 %>%
  ggplot(aes(x = ORB40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$ORB40, y = df_team_pts40$winP, method = "pearson")

#DRB vs win%
df_team_pts40 %>%
  ggplot(aes(x = DRB40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$DRB40, y = df_team_pts40$winP, method = "pearson")

#FG vs win%
df_team_pts40 %>%
  ggplot(aes(x = FG40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$FG40, y = df_team_pts40$winP, method = "pearson")

#FGA vs win%
df_team_pts40 %>%
  ggplot(aes(x = FGA40,y = winP)) + 
  geom_point(colour = "dodgerblue") +
  ylim(0, 100) +
  geom_smooth(method = "lm", colour = "magenta") +
  geom_hline(yintercept = 50, colour = "black", linetype = "dashed")
cor(x = df_team_pts40$FGA40, y = df_team_pts40$winP, method = "pearson")
```

team stats related to pts40
```{r}
#AST40 vs pts40
df_team_pts40 %>%
  ggplot(aes(x = AST40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$AST40, y = df_team_pts40$team_pts40, method = "pearson")

#TOV40 vs pts40
df_team_pts40 %>%
  ggplot(aes(x = TOV40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$TOV40, y = df_team_pts40$team_pts40, method = "pearson")

#STL40 vs pts40
df_team_pts40 %>%
  ggplot(aes(x = STL40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$STL40, y = df_team_pts40$team_pts40, method = "pearson")

#BLK40 vs pts40
df_team_pts40 %>%
  ggplot(aes(x = BLK40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$BLK40, y = df_team_pts40$team_pts40, method = "pearson")

#TRB40 vs pts40
df_team_pts40 %>%
  ggplot(aes(x = TRB40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$TRB40, y = df_team_pts40$team_pts40, method = "pearson")

#DRB vs pts40
df_team_pts40 %>%
  ggplot(aes(x = DRB40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$DRB40, y = df_team_pts40$team_pts40, method = "pearson")

#ORB vs pts40
df_team_pts40 %>%
  ggplot(aes(x = ORB40,y = team_pts40)) + 
  geom_point(colour = "dodgerblue") +
  geom_smooth(method = "lm", colour = "magenta")
cor(x = df_team_pts40$ORB40, y = df_team_pts40$team_pts40, method = "pearson")
```

** choose variables associated with high scoring **
3 explanatory variables associated with win% and total score 40 are:
Points Scored
Assists
Blocks
Total rebounds

**Combine player statistics and player salary data files by *'player names'***
```{r combine files}
# join df_sal & df_pl_stat by "player_name"
df_players <- left_join(x = df_sal, y = df_players, by = c("player_name"))
```


```{r write new file}
write.csv(df_players, file = 'data_processed/2018-19_nba_player-salaries-stats_.csv', row.names = FALSE)
```

Clean processed data file
```{r visualise missing data}
sum(is.na(df_players)) 
naniar::vis_miss(df_players)
#filter out players with no position or game time
df_pl_clean <- drop_na(df_players, Pos)
naniar::vis_miss(df_pl_clean)
```


**Create normalised data by adding variable "points per minute played"**
Explore the data filter out games less than 10 to minimise outliers
```{r}
#ggplot(data = df_players) +
# geom_histogram(mapping = aes(x = salmil), colour = "black", fill = "dodgerblue") +
#  labs(x = "millions", y = "number of players", title = "Player salary distribution per year")

df_players <- mutate(df_players, pts40 = (PTS / MP)* 40)

df_players %>%
  group_by(player_id) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = pts40), colour = "black", fill = "dodgerblue") +
  labs(x = "points per minutes played", y = "number of players", title = "Distribution of points scored per minute played", subtitle = "(games played > 10)")

df_players %>%
  group_by(player_id) %>%
  filter(G > 10) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = pts40), colour = "black", fill = "dodgerblue") +
  labs(x = "points per minutes played", y = "number of players", title = "Distribution of points scored per minute played", subtitle = "(games played > 10)")

```

merge rows, add column sal.mil
```{r}
df_players <- mutate(df_players, salmil = salary / 100000)
```



```{r}
#combine rows with same player names
df_players <- bind_rows(df_players) %>%
   group_by(player_id, player_name, Pos) %>%
   summarise(salmil = mean(salmil, na.rm = TRUE),
              G = sum(G, na.rm = TRUE),
             MP = sum(MP, na.rm = TRUE),
            TRB = sum(TRB, na.rm = TRUE),
            AST = sum(AST, na.rm = TRUE),
            BLK = sum(BLK, na.rm = TRUE),
            PTS = sum(PTS, na.rm = TRUE))
              ungroup(df_players)

df_players40 <- bind_rows(df_players) %>%
   group_by(player_id, player_name) %>%
   summarise(salmil = mean(salmil, na.rm = TRUE),
              G = sum(G, na.rm = TRUE),
             MP = sum(MP, na.rm = TRUE),
            TRB = sum(TRB, na.rm = TRUE),
            AST = sum(AST, na.rm = TRUE),
            BLK = sum(BLK, na.rm = TRUE),
            PTS = sum(PTS, na.rm = TRUE))

df_players <- df_players %>%
  mutate(pts40_ind = (PTS / MP)* 40,
         AST40 = (AST / MP)* 40,
         BLK40 = (BLK / MP)* 40,
         TRB40 = (TRB / MP)* 40)

df_players40 <- df_players40 %>%
  mutate(pts40_ind = (PTS / MP)* 40,
         AST40 = (AST / MP)* 40,
         BLK40 = (BLK / MP)* 40,
         TRB40 = (TRB / MP)* 40)

#exclude data from score value zero
df_players <- df_players[!is.na(df_players$pts40_ind), ]
#eclude players who haven't played a game
df_players <- df_players[df_players$G != 0, ]
```
Remove all missing data rows


look for relationships b/t Player positions and Explanatory Variables
```{r look for correlations}

#player group by pts_ind
df_players %>%
  filter(G > 10) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = reorder(Pos, pts40_ind, FUN = median),
                             y = pts40_ind, colour = reorder(Pos, pts40_ind,
                                                              FUN = median))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))

#player group by AST
df_players %>%
  filter(G > 10) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = reorder(Pos, AST40, FUN = median),
                             y = AST40, colour = reorder(Pos, AST40,
                                                              FUN = median))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))

#player group by BLK
df_players %>%
  filter(G > 10) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = reorder(Pos, BLK40, FUN = median),
                             y = BLK40, colour = reorder(Pos, BLK40,
                                                              FUN = median))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))

#player group by TRB
df_players %>%
  filter(G > 10) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = reorder(Pos, TRB40, FUN = median),
                             y = TRB40, colour = reorder(Pos, TRB40,
                                                              FUN = median))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))
```

```{r multiple linear regreession, team}
fit <- lm(winP ~ team_pts40 + AST40 + BLK40 + TRB40, data = df_team_pts40)
tidy(fit, conf.int = TRUE)
summary(fit)
```

```{r multiple linear regression indiv attack}
fit_ind_att <- lm(salmil ~ AST40 + pts40_ind, data = df_players40, subset = G > 10)
tidy(fit_ind_att, conf.int = TRUE)
summary(fit_ind_att)
```

```{r multiple linear regression indiv defence}
fit_ind_def <- lm(salmil ~ BLK40 + TRB40 + pts40_ind, data = df_players40, subset = G > 10)
tidy(fit_ind_def, conf.int = TRUE)
summary(fit_ind_def)
```


** Test Assumptions **
1. Is the data continuous?
  - the response and explanatory variables for all three models are continuous
2. Independence: 
  - potentially fails in the 1st model (expected wins) due to repeated measures of positions
  - Results are 1.9 for each of the other models which is close to the recommended '2' to ensure independence.
```{r Independence}
#team win %
car::durbinWatsonTest(fit)
#attacking variables
car::durbinWatsonTest(fit_ind_att)
#defensive variables
car::durbinWatsonTest(fit_ind_def)
```
4. Linear relationship
  - there is a very strong apparent relationship between winning % and Team points which is to be expected. The relationships are less significant with the other variables, however previous exploratory observations identified these variables as being useful in the models to determine correlation with salary.
```{r linear relatiohsips}
#team win %
car::avPlots(fit)
#attacking variables
car::avPlots(fit_ind_att)
#defensive variables
car::avPlots(fit_ind_def)
```


5. Detect Outliers:
  - no outliers sit outside 2.5 SD for the team win % and points model
  - several potential outliers exist in the attacking and defensive models however after observing hte raw data points identified around 3SD from the mean i am satisfied in accepting these as true data reflecting player specialty attributes in these measurement variables.
```{r outliers}
#team win %
std_res <- rstandard(fit)
points <- 1:length(std_res)
res_labels <- if_else(abs(std_res) >= 2.5, paste(points), "")

ggplot(data = NULL, aes(x = points, y = std_res)) + geom_point() +
geom_text(aes(label = res_labels), nudge_y = 0.3) + ylim(c(-4,4)) +
geom_hline(yintercept = c(-2.5, 2.5), colour = "red", linetype = "dashed")

#attacking variables
std_res_att <- rstandard(fit_ind_att)
points_att <- 1:length(std_res_att)
res_labels_att <- if_else(abs(std_res_att) >= 3, paste(points_att), "")

ggplot(data = NULL, aes(x = points_att, y = std_res_att)) + geom_point() +
geom_text(aes(label = res_labels_att), nudge_y = 0.3) + ylim(c(-4,4)) +
geom_hline(yintercept = c(-3, 3), colour = "red", linetype = "dashed")

#defensive variables
std_res_def <- rstandard(fit_ind_def)
points_def <- 1:length(std_res_def)
res_labels_def <- if_else(abs(std_res_def) >= 3, paste(points_def), "")

ggplot(data = NULL, aes(x = points_def, y = std_res_def)) + geom_point() +
geom_text(aes(label = res_labels_def), nudge_y = 0.3) + ylim(c(-4.2,4.2)) +
geom_hline(yintercept = c(-3, 3), colour = "red", linetype = "dashed")
```
5. Influential points
```{r cooks distance}
#team win %
cook <- cooks.distance(fit)
cook_labels <- if_else(cook >= 0.2, paste(points), "")

ggplot(data = NULL, aes(x = points, y = cook)) + geom_point() +
geom_text(aes(label = cook_labels), nudge_y = 0.01)

#attacking variables
cook <- cooks.distance(fit_ind_att)
cook_labels <- if_else(cook >= 0.04, paste(points_att), "")

ggplot(data = NULL, aes(x = points_att, y = cook)) + geom_point() +
geom_text(aes(label = cook_labels), nudge_y = 0.01)
#defensive variables
cook <- cooks.distance(fit_ind_def)
cook_labels <- if_else(cook >= 0.04, paste(points_def), "")

ggplot(data = NULL, aes(x = points_def, y = cook)) + geom_point() +
geom_text(aes(label = cook_labels), nudge_y = 0.01)
```
```{r remove influenctial points}
outliers <- c(23, 28, 29)
filtered_team_df <- df_team_pts40 %>%
  filter(!Rk %in% outliers)

fit2 <- fit <- lm(winP ~ team_pts40 + AST40 + BLK40 + TRB40, data = filtered_team_df)
tidy(fit2, conf.int = TRUE)
summary(fit2)
```
** Having looked closely the outliers from the team win percentage model i trust that the data is true team statistic based on real game performances and potentially reflects team styles, therefore will leave this data in the model.

6. Homoscedasticity
```{r hemoscedasticity}
#team win %
res <- residuals(fit)
fitted <- predict(fit)

ggplot(data = NULL, aes(x = fitted, y = res)) +
  geom_point(colour = "dodgerblue") +
  geom_smooth(se = FALSE, colour = "magenta")

#attacking variables
res <- residuals(fit_ind_att)
fitted <- predict(fit_ind_att)

ggplot(data = NULL, aes(x = fitted, y = res)) +
  geom_point(colour = "dodgerblue") +
  geom_smooth(se = FALSE, colour = "magenta")

#defensive variables
res <- residuals(fit_ind_def)
fitted <- predict(fit_ind_def)

ggplot(data = NULL, aes(x = fitted, y = res)) +
  geom_point(colour = "dodgerblue") +
  geom_smooth(se = FALSE, colour = "magenta")
```
Homoscedasticity is maintained for the team win percentage model, yet is present in both individual player models



7. Normally distributed residuals
```{r}
ggplot(data = NULL, aes(sample = res)) + stat_qq() + stat_qq_line()

ggplot(data = NULL, aes(x = res)) + geom_histogram(colour = "black", fill = "dodgerblue")
```
The data is slightly skewed to the left and tailing to the right, however having investigated the data i am comfortable with the underlying context

8. Multicollinearity
```{r}
#team
pairs(formula = ~ winP + team_pts40 + AST40 + BLK40 + TRB40, data = df_team_pts40)
sqrt(car::vif(fit))
#attacking variables
pairs(formula = ~ salmil + pts40_ind + AST40, data = df_players, subset = G > 10)
sqrt(car::vif(fit_ind_att))
#defensive variables
pairs(formula = ~ salmil + pts40_ind + BLK40 + TRB40, data = df_players, subset = G > 10)
sqrt(car::vif(fit_ind_def))
```


```{r}
new_dat <- df_team_pts40 %>% 
  select(Team, winP, AST40, BLK40, TRB40, team_pts40)

fit <- lm(winP ~ AST40 + BLK40 + TRB40 + team_pts40, data = df_team_pts40)

new_dat <- mutate(new_dat, exp_win_perc = predict(fit, newdata = new_dat))

ggplot(new_dat, aes(winP, exp_win_perc, label = Team)) + geom_point(colour = "dodgerblue") +
geom_text(nudge_x = 0.1, cex = 3) +
geom_abline(linetype = "dashed", colour = "magenta")

#indiv attack
df_players <- ungroup(df_players)

fit_ind_att <- lm(salmil ~ AST40 + pts40_ind, data = df_players)

df_players <- mutate(df_players, exp_salmil_att = predict(fit_ind_att, newdata = df_players))

df_players %>%
  filter(G > 10, !Pos %in% c("C", "PF", "C-PF", "PF-C")) %>%
  ggplot () +
  geom_point(mapping = aes(x = exp_salmil_att, y = salmil, colour = Pos)) +
#geom_text(nudge_x = 0.1, cex = 3) +
geom_abline(linetype = "dashed", colour = "magenta")

#indiv defence

fit_ind_def <- lm(salmil ~ BLK40 + TRB40 + pts40_ind, data = df_players)

df_players <- mutate(df_players, exp_salmil_def = predict(fit_ind_def, newdata = df_players))

df_players %>%
  filter(G > 10, !Pos %in% c("SG", "SF", "PG", "SF-SG", "SG-SF")) %>%
  ggplot () +
  geom_point(mapping = aes(x = exp_salmil_def, y = salmil, colour = Pos)) +
#geom_text(nudge_x = 0.1, cex = 3) +
geom_abline(linetype = "dashed", colour = "magenta")
```
```{r}
fit_ind_def <- lm(salmil ~ BLK40 + TRB40 + pts40_ind, data = df_players40)

new_dat_ind <- mutate(new_dat_ind, ext_salmil = predict(fit_ind_def, newdata = new_dat_ind))

```





```{r}
#TRB
df_players %>%
filter(G > 10, Pos %in% c("C", "PF", "PG", "SF", "SG")) %>% 
ggplot() +
geom_histogram(aes(x = TRB40, fill = Pos), colour = "black", bins = 40) + facet_wrap(~Pos, nrow = 5)

#assist
df_players %>%
filter(G > 10, Pos %in% c("C", "PF", "PG", "SF", "SG")) %>% 
ggplot() +
geom_histogram(aes(x = AST40, fill = Pos), colour = "black", bins = 40) + facet_wrap(~Pos, nrow = 5)

#pts40
df_players %>%
filter(G > 10, Pos %in% c("C", "PF", "PG", "SF", "SG")) %>% 
ggplot() +
geom_histogram(aes(x = pts40_ind, fill = Pos), colour = "black", bins = 40) + facet_wrap(~Pos, nrow = 5)
# create panels by match_outcome
```

```{r}
df_players %>%
  filter(G > 10) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = reorder(Pos, AST40, FUN = median),
                             y = AST40, colour = reorder(Pos, AST40,
                                                              FUN = median))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))

#salary by points per minutes played
df_players40 %>% filter(G > 10, PTS > 0) %>% group_by(player_id) %>%
  ggplot(mapping = aes(x = salmil, y = pts40_ind)) + 
  geom_point()

#salary by points per minutes played
df_players %>% filter(G > 10) %>% group_by(player_id) %>%
  ggplot(mapping = aes(x = salmil, y = pts40_ind, colour = Pos)) + 
  geom_point()

#player group by salmil
df_players %>%
  filter(G > 10) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = reorder(Pos, salmil, FUN = median),
                             y = salmil, colour = reorder(Pos, salmil,
                                                              FUN = median))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45))

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
